---
title: "Normal model with nonidentifiable variance"
author: 
- "Subhash Lele"
- "C George Glen"
- "Jose Miguel Ponciano"
date: "`r Sys.Date()`"
output: html_document
---

## Normal model with nonidentifiable variance
First we load the necessary function libraries
```{r message=FALSE, warning=FALSE}
#### ---------------------------------------------------------------------------
#### Load functions



# functions
tausigsq <- function(N, sigsq, tausq) (sigsq*tausq)/( tausq + sigsq )
sdtausq <- function(param,psi.fix){
  tausq = exp(param[[2]])**2
  sdsq  = exp(param[[3]])**2
  psi   = (tausq+sdsq)-psi.fix
  return(psi)
}

profile.analytical.TMB = function(par, obj, fpsi, psi.grid,...){
  lplike.out = matrix(0,length(psi.grid), 2)
  for (i in 1:length(psi.grid)){
    tmp = suppressWarnings(alabama::auglag(par=par, fn = obj$fn, gr = obj$gr,
                                           heq = fpsi, psi.fix=psi.grid[i],
                                           control.outer =list(trace = F)))
    lplike.out[i,] = c(psi.grid[i],tmp$val)
  }
  lplike.out = data.frame(psi=lplike.out[,1],profile.like=lplike.out[,2])
  return(lplike.out)
}

norm_sigmatausqPDF <- function(Y, mu1, sigtausq){
  #llik <- sum(dnorm( Y, mu, sqrt(sigtausq), log = T ))
  lvec <- rep(NA, length( Y ))
  for( i in 1:length( Y ) ){
    lvec <- 1/( sqrt( 2*pi*(sigtausq) ) ) * 
      exp( -( (( Y - mu1 )**2) / (2*( sigtausq )) ) ) 
  }
  llik  <- sum(log(lvec))
  return(llik)
}

norm_tausqPDF <- function(par, Y, tausq){
  mu1       <- par[1]
  sigmasq   <- par[2]**2
  # llik      <- sum(dnorm( Y, mu1, sqrt(sigmasq) + sqrt(tausq), log = T ))
  
  lvec <- rep(NA, length( Y ))
  for( i in 1:length( Y ) ){
    lvec <- 1/( sqrt( 2*pi*(sigmasq + tausq) ) ) * 
      exp( -( (( Y - mu1 )**2) / (2*( sigmasq + tausq )) ) ) 
  }
  llik  <- sum(log(lvec))
  return(-llik)
}


main.nonident <- function(N=3, search.initial=T){
  
  # save results in a list
  res <- list()
  
  for(j in 1:N){
    repeat{
      repeat{
        test.TMB1 <- tryCatch(
          {
            suppressWarnings( run.tmb(profile.par.name="lntau1", 
                                      search.initial = search.initial, n.search = 500,
                                      sim.data=sim.data, model="Norm.nonident.var", 
                                      method.ci="useconfint",
                                      ytol=10, init = list(par=0, hidden=0.1)) )
            
          }, error = function(x){NULL})
        if( !is.null(test.TMB1) ) break
      }
      if( test.TMB1$convergence$pdHess.check == "converged" ) break 
    }
    tmb.res   <- sdreport(test.TMB1$fun.obj)
    
    # summary( tmb.res, p.value = T ) |> tail()
    # test.TMB1$convergence; 
    # test.TMB1$results
    
    res[[j]] <- list( tmb.out=test.TMB1)
  }
  
  
  ### identifiable parameter
  ident.par <- (parms.err.vars1$sd1**2+parms.err.vars1$tau1**2)
  
  ### function to compute the pCI for sdtau using alabama for constrained optimization
  bracket <- 5
  
  #### compute the profile using auglag
  psi.grid.tmb <- seq( 0.001, ident.par+bracket, length.out=100 )
  ana.profile.sdtau <- profile.analytical.TMB(par=c(0,0,0),
                                              obj=test.TMB1$fun.obj, fpsi=sdtausq,
                                              psi.grid=psi.grid.tmb,
                                              control.outer = list(method="BFGS", itmax=5000, NMinit=1))
  ana.profile.sdtau$profile.like  <- exp(-ana.profile.sdtau$profile.like + as.numeric(min(ana.profile.sdtau$profile.like)))
  
  ### draw the unconditional likelihood of Y ~ N(mu, sigma + tau)
  prof.par.sigmatausq <- seq( 0.001, sqrt(ident.par)+bracket, length.out=10000 )
  LLvec.sigmatausq    <- as.numeric()
  for(i in seq_along(prof.par.sigmatausq)){
    LLvec.sigmatausq[i] <- norm_sigmatausqPDF(Y=sim.data[,2],
                                              mu=parms.err.vars1$mu1,
                                              sigtausq = prof.par.sigmatausq[i])
  }
  prof.like.sigmatausq <- exp(LLvec.sigmatausq - max(LLvec.sigmatausq))
  
  ### draw the true profile for tausq 
  prof.par.tausq <- seq(0, ident.par+bracket, by=0.01)
  LLvec.tausq    <- as.numeric() 
  for(i in seq_along(prof.par.tausq)){
    LLvec.tausq[i] <- -optim( fn=norm_tausqPDF, Y=sim.data[,2], 
                              par = c(0.1,1), tausq=prof.par.tausq[i])$value
  }
  LLvec.tausq <- unlist(LLvec.tausq)
  prof.like.tausq <- exp(LLvec.tausq - max(LLvec.tausq))
  
  ### final model results
  supp.res=list(analytical.out=ana.profile.sdtau, 
                ident.par=ident.par, 
                prof.par.sigmatausq=prof.par.sigmatausq, 
                prof.like.sigmatausq=prof.like.sigmatausq,
                prof.par.tausq=prof.par.tausq, 
                prof.like.tausq=prof.like.tausq,
                true.parms=parms.err.vars1)
  res.out <- list( TMBres=res, Ares=supp.res)
  
  return(res.out)
}

### source files
source("TestingLaplaceTMB.functions7.0.R")
source("TestingLaplaceTMB.models2.0.R")
source("TestingLaplaceToolkit4.0.R")

#### ---------------------------------------------------------------------------
#### Load libraries
library(ggplot2)
library(patchwork)
library(knitr)
library(boot)
```

```{r echo=TRUE,  eval=FALSE}

#### ---------------------------------------------------------------------------
#### Generate data
## normal model with non-identifiable variance components
n               <- 100
parms.err.vars1 <- list(mu1=1, sd1=sqrt(4.3), tau1=sqrt(0.9))
fulldf          <- sim.example(n=n, parm.list=parms.err.vars1, model="Norm.nonident.var")
sim.data <- fulldf[,c("X1","Y")]




#### ---------------------------------------------------------------------------
N <- 50
out <- main.nonident(N=N, search.initial=T)
lapply(1:N, function(x) out$TMBres[[x]]$tmb.out$convergence$pdHess.check ) |> unlist()
lapply(1:N, function(x) diagnose(fit=out$TMBres[[x]]$tmb.out)) |> unlist()

out$TMBres[[1]]$tmb.out$results
out$TMBres[[1]]$tmb.out$convergence

## check if anything looks odd
diagnose(out$TMBres[[1]]$tmb.out)



#### ---------------------------------------------------------------------------
### Plot
keep <- which(lapply(1:N, function(x) out$TMBres[[x]]$tmb.out$convergence$pdHess.check ) |> unlist() == "converged")
keep.out <- sapply(keep, function(x) out$TMBres[[x]])


keep.plot <- sapply(c(12,13,18,22), function(x) out$TMBres[x])
plot.res <- list( plot.tmb=keep.plot, Ares=out$Ares  )
# save(plot.res, file="Fig5Data.RData")



N <- length(out)
par(oma=c(1,1,1,1), mar=c(5,5,2,4), mfrow=c(3,3))
#out <- keep.out

for( j in 1:length(keep)){
  
  i <- keep[j]; print(i)
  #i <- j
  
  plot( exp(out$TMBres[[i]]$tmb.out$ci.obj$data$lntau1)**2, 
        out$TMBres[[i]]$tmb.out$ci.obj$data$rel.like, 
        type="l", lwd=2, xlim=c(0, 12), 
        xlab=bquote("Values of" ~ tau^2), ylab="Relative Profile Likelihood", bty="l",
        cex.lab=1.5);
  
  rect(xleft = 0, xright = out$Ares$ident.par, 
       ybottom = 0, ytop =1, 
       border = NA, col = adjustcolor("blue", alpha = 0.1))
  
  lines( out$Ares$analytical.out$psi, out$Ares$analytical.out$profile.like, 
         type="l", col="blue", lty=2, lwd=2);
  
  segments(y0=exp(-qchisq(0.95,1)/2), y1 =exp(-qchisq(0.95,1)/2),
           x0 = min(out$Ares$analytical.out$psi[out$Ares$analytical.out$profile.like > exp(-qchisq(0.95,1)/2)]),
           x1 = max(out$Ares$analytical.out$psi[out$Ares$analytical.out$profile.like > exp(-qchisq(0.95,1)/2)]),
           lwd=2, col = "grey40", lty = par("lty"))
  
  lines( out$Ares$prof.par.sigmatausq, out$Ares$prof.like.sigmatausq, type="l", col="red", lty=3, lwd=2)
  lines( out$Ares$prof.par.tausq, out$Ares$prof.like.tausq, type="l", col="darkgreen", lty=3, lwd=2)
  
  abline( v= (out$Ares$true.parms$sd1**2+out$Ares$true.parms$tau1**2), lty=4, col="black" )
  abline(v=out$Ares$true.parms$tau1**2, lty=5)
}


legend("topright", legend = c( expression("TMB profile for" ~ tau**2), 
                               expression("Derived TMB profile for" ~ tau**2 + sigma**2),
                               expression("True profile for" ~ tau**2 + sigma**2),
                               expression("True profile for" ~ tau**2),
                               expression("True value for" ~ tau**2),
                               expression("ML estimate for" ~ tau**2 + sigma**2),
                               expression("95% Wald-type CI for" ~ tau**2 + sigma**2)),
       pch = c(NA, NA, NA), lty = c(1,2,3,3,4,5,1), lwd = c(2,2,2,2,2,2,3), inset=c(0,0),
       col = c("black","blue","red","darkgreen","black","black","grey40"), bty="n", cex=0.8, y.intersp=0.8)





## plot using ggplot
library(ggplot2)
lt=c("solid", "dashed", "dotted", "dotdash", "longdash", "twodash", "F1")

i <- 3
NIDVarPlot.data <- out$TMBres[[i]]$tmb.out$ci.obj$data
NIDVarPlot <- ggplot(NIDVarPlot.data, aes(exp(lntau1)**2, rel.like) ) + 
  
  geom_rect(aes( xmin = 0, xmax = out$Ares$ident.par, ymin = 0, ymax =1), 
            fill = "grey80", col = NA, alpha=0.1 ) +
  
  geom_line(linewidth=1, aes(linetype ="lineTMBtau"), color="black") +
  geom_line(data=out$Ares$analytical.out, 
            aes(x=psi, y=profile.like, linetype ="lineTMBtausigma"),
            linewidth=1, color="black") +
  geom_line(data=data.frame(x=out$Ares$prof.par.sigmatausq, 
                            y=out$Ares$prof.like.sigmatausq), 
            aes(x=x, y=y, linetype ="lineTrueTauSigma"),
            linewidth=1, color="black") +
  geom_line(data=data.frame(x=out$Ares$prof.par.tausq, 
                            y=out$Ares$prof.like.tausq), 
            aes(x=x, y=y, linetype ="lineTrueTau"),
            linewidth=1, color="black") +
  geom_vline(linewidth=0.95, color = "black",
             mapping = aes(linetype = "TrueTauSigma", 
                           xintercept = out$Ares$true.parms$sd1**2+out$Ares$true.parms$tau1**2),
             show.legend =TRUE, key_glyph = "path") +
  geom_vline(linewidth=0.95, color = "black", 
             mapping = aes(linetype = "TrueTau", xintercept = out$Ares$true.parms$tau1**2),
             show.legend =TRUE, key_glyph = "path") +
  
  geom_segment(aes( y=exp(-qchisq(0.95,1)/2), 
                    yend =exp(-qchisq(0.95,1)/2),
                    x = min(out$Ares$analytical.out$psi[out$Ares$analytical.out$profile.like > exp(-qchisq(0.95,1)/2)]),
                    xend = max(out$Ares$analytical.out$psi[out$Ares$analytical.out$profile.like > exp(-qchisq(0.95,1)/2)]),
                    linetype = "CI95"), lwd=2) +
  xlim(0,10) +
  labs( x=bquote("Values of" ~ tau^2), y="Relative Likelihood", title = "") +
  scale_linetype_manual(name = "", 
                        values = c(lineTMBtau = 4, lineTMBtausigma=2, 
                                   lineTrueTauSigma = 3, lineTrueTau = 1, 
                                   TrueTauSigma = 5, TrueTau = 6, 
                                   CI95 = 7), 
                        labels = c(lineTMBtau = expression("TMB profile for" ~ tau**2), 
                                   lineTMBtausigma=expression("Derived TMB profile for" ~ tau**2 + sigma**2), 
                                   lineTrueTauSigma = expression("True profile for" ~ tau**2 + sigma**2), 
                                   lineTrueTau = expression("True profile for" ~ tau**2), 
                                   TrueTauSigma = expression("ML estimate for" ~ tau**2 + sigma**2), 
                                   TrueTau = expression("True value for" ~ tau**2),
                                   CI95 = expression("95% Wald-type CI for" ~ tau**2 + sigma**2)), 
                        limits = c("lineTMBtau","lineTMBtausigma",
                                   "lineTrueTauSigma", "lineTrueTau", 
                                   "TrueTauSigma", "TrueTau","CI95"), 
                        guide = guide_legend(override.aes = list(color = rep("black",7), 
                                                                 linetype = c(lt[4], lt[2], lt[3], lt[1], 
                                                                              lt[5], lt[6], lt[7]), 
                                                                 linewidth = c(1,1,1,1,1,1,2) )) ) +
  theme_classic() + 
  theme(text = element_text(size=15),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        legend.key.width = unit(4, "line"),        
        axis.ticks.length=unit(0.2, "cm"), legend.position = "bottom") 
NIDVarPlot
#ggsave(filename="Figure5.png", plot=NIDVarPlot, width = 15, height=9)  


```


Remove previous object file and DLL if present
```{r}
#### ---------------------------------------------------------------------------
dyn.unload(dynlib(model_name))

if (length(list.files(pattern = "\\.o$", full.names = TRUE))>0) {
  file.remove(list.files(pattern = "\\.o$", full.names = TRUE))
}
if (length(list.files(pattern = "\\.so$", full.names = TRUE))>0) {
  file.remove(list.files(pattern = "\\.so$", full.names = TRUE))
}
if (length(list.files(pattern = "\\.cpp$", full.names = TRUE))>0) {
  file.remove(list.files(pattern = "\\.cpp$", full.names = TRUE))
}
#