---
title: "Poisson-Normal Squared model"
author: 
- "Subhash Lele"
- "C George Glen"
- "Jose Miguel Ponciano"
date: "`r Sys.Date()`"
output: html_document
---

## Poisson-Normal Squared, one parameter
First we upload the necessary function libraries
```{r message=FALSE, warning=FALSE}
#### ---------------------------------------------------------------------------
#### Load functions
source("TestingLaplaceToolkit4.0.R")
source("TestingLaplaceTMB.functions7.0.R") 
source("TestingLaplaceTMB.models2.0.R") 

#### ---------------------------------------------------------------------------
#### Load libraries
library(ggplot2)
library(patchwork)
library(knitr)
library(boot)
```

```{r echo=TRUE, eval=FALSE}

#### ---------------------------------------------------------------------------
# Simulating from Model 3: Latent: X ~ Normal(mu, sd1), U = X^2 (~Chi-square(1))
# Observation model: Y ~ Poisson(lambda=U)
parms.Usq <- list(mu=-1, sd1=5)
out.simUsq <- sim.example(n=300, parm.list=parms.Usq, model="Normsq.Pois")


#### ---------------------------------------------------------------------------
# DC run to estimate beta assuming all other parameters are known
ModUsq.dc.beta <- onedcfitrun(model="Normsq.Pois", true.parms=parms.Usq,
                              sim.data = out.simUsq,n.iter=5000,n.adapt=100,
                              n.update=100,thin=1, n.chains=4,clones.seq=c(1,2,8,16))


# kable(round(dcdiag(ModUsq.dc.beta),digits=4), caption="Diagnostics for Data Cloning runs using multiple clones to assess convergence")

Beta.hat <- summary(ModUsq.dc.beta)[[1]][1]
# Or simply use 
Beta.hatUsq.CI <- confint(ModUsq.dc.beta)



#### ---------------------------------------------------------------------------
# True MC-profile likelihood (i.e. computing knowing true parameter
# values for the distribution of the hidden process)
# Model 3: X~Normal, Y~Poisson(X^2)
MCprof.ModUsq <- MCprofile1d(B=150000, parm.list=parms.Usq, sim.data=out.simUsq, 
                             bracket=8,prec.prof=0.02, model="Normsq.Pois",
                             plot.it=FALSE,overlay=FALSE)


#### ---------------------------------------------------------------------------
# LOESS smoothing of MC profile
psi.grid <-  seq(Beta.hat-6,Beta.hat+10,0.01)
MCdata <- data.frame(beta_values=MCprof.ModUsq[,1],RelPL=MCprof.ModUsq[,2])
tmp <- loess(RelPL ~ beta_values,data=MCdata, span=0.08)
newdata.4Mod <- data.frame(beta_values=psi.grid)
MCprof.smooth <- predict(tmp,newdata=newdata.4Mod)
where.nas <- which(is.na(MCprof.smooth), arr.ind=TRUE)
MCprof.plot <- MCprof.smooth[-where.nas]/max(MCprof.smooth[-where.nas])
psi.grid.plot <- psi.grid[-where.nas]
gridmcdf <- data.frame(Beta.values=psi.grid.plot, Rel.Prof.Like = MCprof.plot)



#### ---------------------------------------------------------------------------
## TMB
parms4tmb <- list(mu=parms.Usq$mu, lnsd1 = log(parms.Usq$sd1))

testUsq <- run.tmb(profile.par.name="mu", profile.par=MCprof.ModUsq[,1], 
                   known.parm=parms4tmb, sim.data=out.simUsq, model="Normsq.Pois")



#### ---------------------------------------------------------------------------
# Plot the MC profile with the TMB profile
PNsq1 <- ggplot(gridmcdf, aes(Beta.values, Rel.Prof.Like) ) + 
  geom_line(linewidth=1, aes(linetype ="black.line"), color="black") + xlim( c(-10,8) ) +
  labs( x=bquote("Values of" ~ beta), y="Relative Likelihood", title = "A") +
  geom_line(data = testUsq$prof, aes( x=param, y=rel.like, linetype="red.line"), color = "red", linewidth=1) +
  
  geom_line(data=data.frame(x=as.vector(Beta.hatUsq.CI), y=c(0,0)), aes(x=x,y=y, linetype = "blue.line"), color = "blue", linewidth=2)+
  geom_vline(linetype="dotted",linewidth=0.95, mapping = aes(linetype = "black.line2", xintercept = Beta.hat), color = "black")+
  
  geom_hline(linetype="dotted",linewidth=0.9, mapping = aes(linetype="gray.line", yintercept = exp(-qchisq(p=0.95,df=1)/2)), color = "gray")+ 
  scale_linetype_manual(name = "", values = c(red.line = 2, black.line = 1, blue.line = 1, black.line2 = 1, gray.line =2), 
                        labels = c(red.line = "TMB", black.line = "MC", blue.line = "DC-CI", black.line2 = "MLE", gray.line = bquote("exp(-0.5"~ chi[0.95]^2~")")), 
                        limits = c("red.line", "black.line", "blue.line", "black.line2", "gray.line"), 
                        guide = guide_legend(override.aes = list(color = c("red", "black", "blue", "black", "gray"), linetype = c("dotted","solid","solid",    "dotted","dotted"), linewidth = c(1, 1,1, 1,1) )) ) +
  theme_classic() + 
  theme(text = element_text(size=15),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.ticks.length=unit(0.2, "cm"), legend.position = "bottom") 
PNsq1

```



b) Comparison of the MC likelihood with GT:
```{r echo=TRUE, eval=FALSE}


#### ---------------------------------------------------------------------------
parms.Usq$mu <- Beta.hat

# Sampling from the conditional unobserved process
Xpost.ModUsq <- Xpost.samp(model="Normsq.Pois", 
                           parm.list=parms.Usq,
                           sim.data=out.simUsq, n.iter=50100,n.adapt=1000,
                           n.update=1000,thin=4,n.chains=5)

Xpost.binned <- do.call(rbind,Xpost.ModUsq)


B <- nrow(Xpost.binned)

XpostSample <- Xpost.binned

sd1 <- parms.Usq$sd1
beta<- parms.Usq$mu
n <- nrow(out.simUsq)
Y <- out.simUsq[,1]

beta <- Beta.hat
bracket <- 6



#### ---------------------------------------------------------------------------
# Computing GT's denominator for every sampled X and every value of the
# parameter 'beta' we are profiling over
prec.prof <- 0.02
beta.profs <- seq(from=beta-bracket,to=beta+bracket, by=prec.prof)
len.prof <- length(beta.profs)
prof.vec <- rep(0,len.prof)

for(i in 1:len.prof){
  
  ith.beta <- beta.profs[i]
  GT.logdiff <- rep(0,B)
  
  for(j in 1:B){
    X.star <- XpostSample[j,]
    GT.lnumer.persamp<- sum(dnorm(x=X.star, mean=ith.beta, sd=sd1, log=TRUE))
    #  +dchisq(x=(X.star^2)/(sd1^2), df=1,ncp=(ith.beta^2), log=TRUE)
    # + dpois(x=Y, lambda=Usq.star, log=TRUE) --> it cancels
    
    GT.ldenom.persamp <-  sum(dnorm(x=X.star, mean=beta, sd=sd1, log=TRUE))
    # + dchisq(x=(X.star^2)/(sd1^2), df=1,ncp=(beta^2), log=TRUE)
    # + dpois(x=Y, lambda=Usq.star, log=TRUE) --> it cancels
    
    GT.logdiff[j] <- GT.lnumer.persamp-GT.ldenom.persamp
  }
  
  ratio.prof.vec.star <- exp(GT.logdiff)
  prof.vec[i] <- mean(ratio.prof.vec.star)
  
}

```

When plotted, we see that the GT algorithm gives a peak around the MLE.  The tilting was done using the MLE (-3.849896 here) which is far from the true value.
```{r echo=TRUE, eval=FALSE}


#### ---------------------------------------------------------------------------
# Plotting the MC profile with the GT profile
PNsq2 <- ggplot(gridmcdf, aes(Beta.values, Rel.Prof.Like) ) + 
  geom_line(linewidth=1, aes(linetype ="black.line"), color="black") + xlim( c(-10,8) ) +
  labs( x=bquote("Values of" ~ beta), y="Relative Likelihood", title = "A") +
  geom_line(data = data.frame(beta.profs=beta.profs, prof.vec=prof.vec), aes( x=beta.profs, y=prof.vec, linetype="red.line"), color = "red", linewidth=1) +
  
  geom_line(data=data.frame(x=as.vector(Beta.hatUsq.CI), y=c(0,0)), aes(x=x,y=y, linetype = "blue.line"), color = "blue", linewidth=2)+
  geom_vline(linetype="dotted",linewidth=0.95, mapping = aes(linetype = "black.line2", xintercept = Beta.hat), color = "black")+
  
  geom_hline(linetype="dotted",linewidth=0.9, mapping = aes(linetype="gray.line", yintercept = exp(-qchisq(p=0.95,df=1)/2)), color = "gray")+ 
  scale_linetype_manual(name = "", values = c(red.line = 2, black.line = 1, blue.line = 1, black.line2 = 1, gray.line =2), 
                        labels = c(red.line = "GT", black.line = "MC", blue.line = "DC-CI", black.line2 = "MLE", gray.line = bquote("exp(-0.5"~ chi[0.95]^2~")")), 
                        limits = c("red.line", "black.line", "blue.line", "black.line2", "gray.line"), 
                        guide = guide_legend(override.aes = list(color = c("red", "black", "blue", "black", "gray"), linetype = c("dotted","solid","solid",    "dotted","dotted"), linewidth = c(1, 1,1, 1,1) )) ) +
  theme_classic() + 
  theme(text = element_text(size=15),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.ticks.length=unit(0.2, "cm"), legend.position = "bottom") 



#### ---------------------------------------------------------------------------
PNsq.all <- ggplot(gridmcdf, aes(Beta.values, Rel.Prof.Like) ) + 
  geom_line(linewidth=1, aes(linetype ="black.line"), color="black") + xlim( c(-10,8) ) +
  labs( x=bquote("Values of" ~ beta), y="Relative Likelihood", title = "") +
  geom_line(data = testUsq$prof, aes( x=param, y=rel.like, linetype="red.line"), color = "black", linewidth=1) +
  geom_line(data = data.frame(beta.profs=beta.profs, prof.vec=prof.vec), aes( x=beta.profs, y=prof.vec, linetype="anothered.line"), color = "black", linewidth=1) +
  
  geom_line(data=data.frame(x=as.vector(Beta.hatUsq.CI), y=c(0,0)), aes(x=x,y=y, linetype = "blue.line"), color = "black", linewidth=2)+
  geom_vline(linetype=5,linewidth=0.95, mapping = aes(linetype = "black.line2", xintercept = Beta.hat), color = "black")+
  
  geom_hline(linetype=6,linewidth=0.9, mapping = aes(linetype="gray.line", yintercept = exp(-qchisq(p=0.95,df=1)/2)), color = "black")+ 
  scale_linetype_manual(name = "", values = c(red.line = 4, anothered.line=2, black.line = 3, blue.line = 1, black.line2 = 5, gray.line =6), 
                        labels = c(red.line = "TMB", anothered.line="GT", black.line = "MC", blue.line = "DC-CI", black.line2 = "MLE", gray.line = bquote("exp(-0.5"~ chi[0.95]^2~")")), 
                        limits = c("red.line","anothered.line" ,"black.line", "blue.line", "black.line2", "gray.line"), 
                        guide = guide_legend(override.aes = list(color = rep("black",6), linetype = c(4,2,3,1,5,6), linewidth = c(1,1,1, 2,1,1) )) ) +
  theme_classic() + 
  theme(text = element_text(size=15),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        legend.key.width = unit(4, "line"),        
        axis.ticks.length=unit(0.2, "cm"), legend.position = "bottom") 
# ggsave(filename="Figure3.png", plot=PNsq.all,width = 12, height=8)  


# save(PNsq1v2,file="PNsq1.RData")
# save(PNsq2v2,file="PNsq2.RData")
# save.image("PNsq.RData")


```




Doing the GT algorithm by tilting at the true value does not give a different profile.
```{r echo=TRUE, eval=FALSE}

#### ---------------------------------------------------------------------------
# Tilting from the true value allows GT not to die

parms.Usq$mu <- -1 #I will tilt from the true value


#### ---------------------------------------------------------------------------
# Sampling from the conditional unobserved process
Xpost.ModUsq2 <- Xpost.samp(model="Normsq.Pois", parm.list=parms.Usq,
                            sim.data=out.simUsq, n.iter=50100,n.adapt=1000,
                            n.update=1000,thin=4,n.chains=5)

Xpost.binned2 <- do.call(rbind,Xpost.ModUsq2)

B <- nrow(Xpost.binned)

XpostSample <- Xpost.binned

sd1 <- parms.Usq$sd1
beta<- parms.Usq$mu
n <- nrow(out.simUsq)
Y <- out.simUsq[,1]

beta <- Beta.hat
bracket <- 6


#### ---------------------------------------------------------------------------
# Computing GT's denominator for every sampled X and every value of the
# parameter 'beta' we are profiling over
prec.prof <- 0.02
beta.profs <- seq(from=beta-bracket,to=beta+bracket, by=prec.prof)
len.prof <- length(beta.profs)
prof.vec <- rep(0,len.prof)

for(i in 1:len.prof){
  
  ith.beta <- beta.profs[i]
  GT.logdiff <- rep(0,B)
  
  for(j in 1:B){
    X.star <- XpostSample[j,]
    GT.lnumer.persamp<- sum(dnorm(x=X.star, mean=ith.beta, sd=sd1, log=TRUE))
    #  +dchisq(x=(X.star^2)/(sd1^2), df=1,ncp=(ith.beta^2), log=TRUE)
    # + dpois(x=Y, lambda=Usq.star, log=TRUE) --> it cancels
    
    GT.ldenom.persamp <-  sum(dnorm(x=X.star, mean=beta, sd=sd1, log=TRUE))
    # + dchisq(x=(X.star^2)/(sd1^2), df=1,ncp=(beta^2), log=TRUE)
    # + dpois(x=Y, lambda=Usq.star, log=TRUE) --> it cancels
    
    GT.logdiff[j] <- GT.lnumer.persamp-GT.ldenom.persamp
  }
  
  ratio.prof.vec.star <- exp(GT.logdiff)
  prof.vec[i] <- mean(ratio.prof.vec.star)
  
}

#### ---------------------------------------------------------------------------
# Plotting the MC profile with the GT profile
par(oma=c(1,1,1,1), mar=c(4,5,2,2))
plot(psi.grid.plot,MCprof.plot, type="l", col="black", lwd=2, xlim=c(-10,8), 
     xlab=bquote("Values of" ~ beta), ylab="Relative Likelihood", bty="l", 
     cex.lab=1.5)
points(beta.profs,prof.vec, type="l", lty=2, lwd=2, col="red")
points(x=Beta.hatUsq.CI,y=c(0,0), col="blue", type="l" ,lty=1,lwd=4)
abline(v=Beta.hat, lwd=2,lty=2)
abline(h=exp(-qchisq(p=0.95,df=1)/2),lty=2, col="gray")
legend("topright", legend=c("GT", "MC", "MLE","DC-CI",bquote("exp(-0.5"~ chi[0.95]^2~")")), col=c("red", "black","black", "blue","gray"), lty=c(2,1,2,1,2), lwd=c(2,2,2,4,1), bty="n", cex=1.15)


```



Remove previous object file and DLL if present
```{r}
#### ---------------------------------------------------------------------------
dyn.unload(dynlib(model_name))

if (length(list.files(pattern = "\\.o$", full.names = TRUE))>0) {
  file.remove(list.files(pattern = "\\.o$", full.names = TRUE))
}
if (length(list.files(pattern = "\\.so$", full.names = TRUE))>0) {
  file.remove(list.files(pattern = "\\.so$", full.names = TRUE))
}
if (length(list.files(pattern = "\\.cpp$", full.names = TRUE))>0) {
  file.remove(list.files(pattern = "\\.cpp$", full.names = TRUE))
}
#