---
title: "Probit-Normal model"
author: 
- "Subhash Lele"
- "C George Glen"
- "Jose Miguel Ponciano"
date: "`r Sys.Date()`"
output: html_document
---

## Probit-Normal, one parameter
First we upload the necessary function libraries
```{r message=FALSE, warning=FALSE}
#### ---------------------------------------------------------------------------
#### Load functions
source("TestingLaplaceToolkit4.0.R")
source("TestingLaplaceTMB.functions7.0.R") 
source("TestingLaplaceTMB.models2.0.R") 

#### ---------------------------------------------------------------------------
#### Load libraries
library(ggplot2)
library(patchwork)
library(knitr)
library(boot)
```

a) Comparison of the MC likelihood function with TMB
```{r echo=TRUE,  eval=FALSE}

#### ---------------------------------------------------------------------------
#### Generate data
parms.probit <- list(beta=-1, df1=2, sd1=3)
out.simProbN <- sim.example(n=100, parm.list=parms.probit, model="Probit.Norm")
parms.probit <- parms.err.varsProbN

ModProbN.dc.beta <- onedcfitrun(model="Probit.Norm", true.parms=parms.probit,
                            sim.data = out.simProbN,n.iter=51000,n.adapt=1000,
                            n.update=100,thin=10, n.chains=4,clones.seq=c(1,16,32),
                            parallel=FALSE)



#### ---------------------------------------------------------------------------
# kable(round(dcdiag(ModProbN.dc.beta),digits=4), 
#       caption="Diagnostics for Data Cloning runs using multiple clones to assess convergence")

Beta.hat <- summary(ModProbN.dc.beta)[[1]][1]
# Or simply use 
Beta.hatProbN.CI <- confint(ModProbN.dc.beta)


#### ---------------------------------------------------------------------------
#### MC profile
MCprof.ModProbN <- MCprofile1d(B=50000, parm.list=parms.probit, sim.data=out.simProbN, 
                           bracket=4,prec.prof=0.01, model="Probit.Norm",
                           plot.it=FALSE,overlay=FALSE)


#### ---------------------------------------------------------------------------
# LOESS smoothing of MC profile
psi.grid <-  seq(Beta.hat-3.5,Beta.hat+3.5,0.01)
MCdata <- data.frame(beta_values=MCprof.ModProbN[,1],RelPL=MCprof.ModProbN[,2])
tmp <- loess(RelPL ~ beta_values,data=MCdata, span=0.08)
newdata.4Mod <- data.frame(beta_values=psi.grid)
MCprof.smooth <- predict(tmp,newdata=newdata.4Mod)
where.nas <- which(is.na(MCprof.smooth), arr.ind=TRUE)
if(length(where.nas)>0){
  MCprof.plot <- MCprof.smooth[-where.nas]/max(MCprof.smooth[-where.nas])
  psi.grid.plot <- psi.grid[-where.nas]
}else{
  MCprof.plot <- MCprof.smooth/max(MCprof.smooth)
  psi.grid.plot <- psi.grid

}
gridmcdf <- data.frame(Beta.values=psi.grid.plot, Rel.Prof.Like=MCprof.plot)




#### ---------------------------------------------------------------------------
# TMB profile
parms4tmb <- list(beta=parms.probit$beta, df1=parms.probit$df1,
                  lnsd1 = log(parms.probit$sd1))

testProbN <- run.tmb(profile.par.name="beta", profile.par=MCprof.ModProbN[,1], 
                 known.parm=parms4tmb, sim.data=out.simProbN, model="Probit.Norm")
tmbdataframe <- data.frame(param=testProbN$prof$param, rel.like=testProbN$prof$rel.like)



#### ---------------------------------------------------------------------------
# Plot the MC profile with the TMB profile
ProbitN1 <- ggplot(gridmcdf, aes(Beta.values, Rel.Prof.Like) ) + 
  geom_line(linewidth=1, aes(linetype ="black.line"), color="black") + xlim( c(-3,1) ) +
  labs( x=bquote("Values of" ~ beta), y="Relative Likelihood", title = "A") +
  geom_line(data =tmbdataframe , aes( x=param, y=rel.like, linetype="red.line"), color = "red", linewidth=1) +
   
  geom_line(data=data.frame(x=as.vector(Beta.hatProbN.CI), y=c(0,0)), aes(x=x,y=y, linetype = "blue.line"), color = "blue", linewidth=2)+
  geom_vline(linetype="dotted",linewidth=0.95, mapping = aes(linetype = "black.line2", xintercept = Beta.hat), color = "black")+
 
  geom_hline(linetype="dotted",linewidth=0.9, mapping = aes(linetype="gray.line", yintercept = exp(-qchisq(p=0.95,df=1)/2)), color = "gray")+ 
   scale_linetype_manual(name = "", values = c(red.line = 2, black.line = 1, blue.line = 1, black.line2 = 1, gray.line =2), 
                          labels = c(red.line = "TMB", black.line = "MC", blue.line = "DC-CI", black.line2 = "MLE", gray.line = bquote("exp(-0.5"~ chi[0.95]^2~")")), 
                          limits = c("red.line", "black.line", "blue.line", "black.line2", "gray.line"), 
                          guide = guide_legend(override.aes = list(color = c("red", "black", "blue", "black", "gray"), linetype = c("dotted","solid","solid",    "dotted","dotted"), linewidth = c(1, 1,1, 1,1) )) ) +
  theme_classic() + 
  theme(text = element_text(size=15),
             panel.grid.major = element_blank(),
             panel.grid.minor = element_blank(),
             strip.background = element_blank(),
             axis.ticks.length=unit(0.2, "cm"), legend.position = "bottom") 
ProbitN1

  

```


b) Comparing GT to the MC likelihood
```{r echo=TRUE,  eval=FALSE}


#### ---------------------------------------------------------------------------
# Sampling from the conditional unobserved process
Xsamp.ProbN.par      <- parms.probit
Xsamp.ProbN.par$beta <- Beta.hat

Xpost.ModProbN <- Xpost.samp(model="Probit.Norm", parm.list=Xsamp.ProbN.par,
                         sim.data=out.simProbN, n.iter=51000,n.adapt=1000,
                         n.update=1000,thin=4,n.chains=3)
Xpost.binned <- do.call(rbind,Xpost.ModProbN)


#### ---------------------------------------------------------------------------
# GT likelihood calculation "by hand"
# Model5.GTprof <- GT.prof(model="Probit.Norm", parm.list=parms.probit,sim.data=out.sim5,
#                          XpostSample=Xpost.binned, bracket=4, prec.prof=0.01,
#                          plot.it=FALSE, overlay=FALSE,mycol="blue",
#                          dotted=1)

XpostSample <- Xpost.binned
df1  <- parms.probit$df1
sd1  <- parms.probit$sd1
beta <- Beta.hat
n <- nrow(out.simProbN)
W <- out.simProbN[,1]
Y <- out.simProbN[,2]

B <- nrow(XpostSample)
# Computing GT's denominator for every sampled X and every value of the
# parameter 'beta' we are profiling over

bracket <- 4
prec.prof <- 0.01 
beta.profs <- seq(from=beta-bracket,to=beta+bracket, by=prec.prof)
len.prof <- length(beta.profs)
prof.vec <- rep(0,len.prof)

for(i in 1:len.prof){
  
  ith.beta <- beta.profs[i]
  #GT.lnumer.vec <- rep(0,B)
  #GT.ldenom.vec <- rep(0,B)
  GT.logdiff <- rep(0,B)
  for(j in 1:B){
    X.star <- XpostSample[j,]
    pvec.num   <- pnorm(q=ith.beta*X.star, mean=0, sd=1)
    pvec.denom <- pnorm(q=beta*X.star, mean=0,sd=1)
    pvec.num[pvec.num==1]<- 0.99
    pvec.num[pvec.num==0] <- .Machine$double.xmin
    pvec.denom[pvec.denom==1]<- 0.99
    pvec.denom[pvec.denom==0] <- .Machine$double.xmin
    
    GT.lnumer.persamp <- sum(Y*log(pvec.num) + (1-Y)*log(1-pvec.num)
                             + dnorm(x=W, mean=X.star,sd=sd1, log=TRUE) +
                               dt(x=X.star, df=df1, ncp=0,log=TRUE)) 
    GT.ldenom.persamp <- sum(Y*log(pvec.denom) + (1-Y)*log(1-pvec.denom)
                             + dnorm(x=W, mean=X.star,sd=sd1, log=TRUE) +
                               dt(x=X.star, df=df1, ncp=0,log=TRUE))    
    #GT.lnumer.persamp<- dbinom(x=Y, size=1, prob=1/(1+exp(-ith.beta*X.star)), log=TRUE) 
    #+ dnorm(x=W, mean=X.star,sd=sd1, log=TRUE) +
    #dt(x=X.star, df=df1, ncp=0,log=TRUE))  --> this part cancels!
    
    #GT.ldenom.persamp <-  dbinom(x=Y, size=1, prob=1/(1+exp(-beta*X.star)), log=TRUE)
    
    GT.logdiff[j] <- sum(GT.lnumer.persamp-GT.ldenom.persamp)
  }
  
  ratio.prof.vec.star <- exp(GT.logdiff)
  prof.vec[i] <- mean(ratio.prof.vec.star)
  
}
GTdf <- data.frame(beta.profs=beta.profs, rel.prof=prof.vec)
#beta.max <- beta.profs[prof.vec==max(prof.vec)]
#beta <- beta.max



#### ---------------------------------------------------------------------------
# Plotting the MC profile with the GT profile
ProbitN2 <- ggplot(gridmcdf, aes(Beta.values, Rel.Prof.Like) ) + 
  geom_line(linewidth=1, aes(linetype ="black.line"), color="black") + xlim( c(-3,1) ) +
  labs( x=bquote("Values of" ~ beta), y="Relative Likelihood", title = "B") +
  geom_line(data = GTdf, aes( x=beta.profs, y=rel.prof, linetype="red.line"), color = "red", linewidth=1) +
   
  geom_line(data=data.frame(x=as.vector(Beta.hatProbN.CI), y=c(0,0)), aes(x=x,y=y, linetype = "blue.line"), color = "blue", linewidth=2)+
  geom_vline(linetype="dotted",linewidth=0.95, mapping = aes(linetype = "black.line2", xintercept = Beta.hat), color = "black")+
 
  geom_hline(linetype="dotted",linewidth=0.9, mapping = aes(linetype="gray.line", yintercept = exp(-qchisq(p=0.95,df=1)/2)), color = "gray")+ 
   scale_linetype_manual(name = "", values = c(red.line = 2, black.line = 1, blue.line = 1, black.line2 = 1, gray.line =2), 
                          labels = c(red.line = "GT", black.line = "MC", blue.line = "DC-CI", black.line2 = "MLE", gray.line = bquote("exp(-0.5"~ chi[0.95]^2~")")), 
                          limits = c("red.line", "black.line", "blue.line", "black.line2", "gray.line"), 
                          guide = guide_legend(override.aes = list(color = c("red", "black", "blue", "black", "gray"), linetype = c("dotted","solid","solid",    "dotted","dotted"), linewidth = c(1, 1,1, 1,1) )) ) +
  theme_classic() + 
  theme(text = element_text(size=15),
             panel.grid.major = element_blank(),
             panel.grid.minor = element_blank(),
             strip.background = element_blank(),
             axis.ticks.length=unit(0.2, "cm"), legend.position = "bottom")
ProbitN2





#### ---------------------------------------------------------------------------  
ProbitN.all <- ggplot(gridmcdf, aes(Beta.values, Rel.Prof.Like) ) + 
  geom_line(linewidth=1, aes(linetype ="black.line"), color="black") + xlim( c(-3,1) ) +
  labs( x=bquote("Values of" ~ beta), y="Relative Likelihood", title = "") +
  geom_line(data = tmbdataframe, aes( x=param, y=rel.like, linetype="red.line"), color = "black", linewidth=1) +
  geom_line(data = GTdf, aes( x=beta.profs, y=rel.prof, linetype="anothered.line"), color = "black", linewidth=1) +
     
  geom_line(data=data.frame(x=as.vector(Beta.hatProbN.CI), y=c(0,0)), aes(x=x,y=y, linetype = "blue.line"), color = "black", linewidth=2)+
  geom_vline(linetype=5,linewidth=0.95, mapping = aes(linetype = "black.line2", xintercept =  -0.2149747), color = "black")+
 
  geom_hline(linetype=6,linewidth=0.9, mapping = aes(linetype="gray.line", yintercept = exp(-qchisq(p=0.95,df=1)/2)), color = "black")+ 
   scale_linetype_manual(name = "", values = c(red.line = 4, anothered.line=2, black.line = 3, blue.line = 1, black.line2 = 5, gray.line =6), 
                          labels = c(red.line = "TMB", anothered.line="GT", black.line = "MC", blue.line = "DC-CI", black.line2 = "MLE", gray.line = bquote("exp(-0.5"~ chi[0.95]^2~")")), 
                          limits = c("red.line","anothered.line" ,"black.line", "blue.line", "black.line2", "gray.line"), 
                          guide = guide_legend(override.aes = list(color = rep("black",6), linetype = c(4,2,3,1,5,6), linewidth = c(1,1,1, 2,1,1) )) ) +
  theme_classic() + 
  theme(text = element_text(size=15),
             panel.grid.major = element_blank(),
             panel.grid.minor = element_blank(),
             strip.background = element_blank(),
             legend.key.width = unit(4, "line"),        
             axis.ticks.length=unit(0.2, "cm"), legend.position = "bottom")     
ProbitN.all


# save(ProbitN1,file="ProbitN1v2.RData")
# save(ProbitN2,file="ProbitN2v2.RData")
# save(ProbitN.all, file="ProbitNall.RData")
# save.image("ProbitN-Onepar.RData")


```


Remove previous object file and DLL if present
```{r}
#### ---------------------------------------------------------------------------
dyn.unload(dynlib(model_name))

if (length(list.files(pattern = "\\.o$", full.names = TRUE))>0) {
  file.remove(list.files(pattern = "\\.o$", full.names = TRUE))
}
if (length(list.files(pattern = "\\.so$", full.names = TRUE))>0) {
  file.remove(list.files(pattern = "\\.so$", full.names = TRUE))
}
if (length(list.files(pattern = "\\.cpp$", full.names = TRUE))>0) {
  file.remove(list.files(pattern = "\\.cpp$", full.names = TRUE))
}
#